---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
keep_md=TRUE

```

# Loading and Processing Data

```{r}
#load data 
path<-"~/coursera/repdata_data_activity/activity.csv"
activity<-read.csv(file=path,
colClasses=c("integer", "Date", "integer"))
#remove na
steps<-na.omit(activity)
```

# What is the mean total of steps taken per day?

```{r}

#sum steps by date, calculate mean and median
sum_steps<-aggregate(steps~date, steps, sum)
mean_steps<-aggregate(steps~date, steps, mean)
median_steps<-aggregate(steps~date, steps, median)
```

The mean total of steps taken per day is:
```{r}
mean_steps
```

The median is:
```{r}
median_steps
```


Here is a histogram of the total steps taken per day

```{r}

plot(sum_steps, type="h", main="Total Steps Taken Per
Day", xlab="Date", ylab="Steps", lwd=2)
```


# What is the average daily activity pattern?
Here is a time series plot of the average number of steps taken per 5 minute interval averaged across all days:
```{r}

#create avg steps by interval across all days
mean_interval<-aggregate(steps~interval, 
steps, mean)

#plot line graph and return interval w/max steps
plot(mean_interval, type="l", main=
"Mean Steps Per 5-Sec Interval", 
xlab="Interval", ylab="Steps")

max_interval<-mean_interval$interval[mean_interval$steps==max(mean_interval$steps)]

```

The 5 minute interval with the maximum number of steps is:

```{r}
max_interval
```

# Imputing missing values

```{r}

#calculate total NA's in "activity"
totalNA<-sum(is.na(activity$steps))

```

The total number of NA's in dataset is:

```{r}
totalNA
```

```{r, cache=TRUE}
#replace missing values and recalculate sum, mean and ##median steps, plot new steps/day
library(mice)
newframe<-subset(activity, 
subset=, select=c("steps", "interval"))
imp<-mice(newframe, m=5, method="pmm", seed=500)
newframe2<-complete(imp,1)
Date<-activity$date
new_activity<-cbind(newframe2, Date)
sum_steps2<-aggregate(steps~Date,
                      new_activity, sum)
mean_steps2<-aggregate(steps~Date, 
new_activity, mean)
median_steps2<-aggregate(steps~Date, 
new_activity, median)
```

Here is a histogram of of the total steps taken each day after imputation of missing data:
```{r, echo=TRUE}
plot(sum_steps2, type="h", main="Total Steps/Day 
After Imputation", xlab="Date", ylab="Steps", 
lwd=2)

```

The mean steps taken per day after imputing missing data is:
```{r}
mean_steps2
```
The median steps taken per day after imputating missing data is:
```{r}
median_steps2
```

#Are there differences in patterns between weekdays and weekends?

```{r}
#add new column to note weekday or weekend
weekdays<-weekdays(new_activity$Date, 
                   abbreviate=FALSE)

weekdays<-ifelse(weekdays %in% 
  c("Saturday","Sunday"), "Weekend", weekdays)
weekdays<-ifelse(weekdays%in% c("Monday", 
"Tuesday", "Wednesday", "Thursday", "Friday"), 
"Weekday", weekdays)

new_frame3<-cbind(new_activity, weekdays)

#new data frames for weekdays and weekends and 

MTWRF<-subset(new_frame3, subset=weekdays==
"Weekday",select=c("steps", "interval", 
                   "weekdays"))
SS<-subset(new_frame3, subset=weekdays==
"Weekend", select=c("steps", "interval", 
                    "weekdays"))
newSS<-aggregate(steps~interval, SS, mean)
newMTWRF<-aggregate(steps~interval, MTWRF, mean)
```

Here is a time series plot of the average numer of steps taken per 5 minute interval averaged across weekdays and weekends:
```{r, echo=TRUE}
#plot mean steps by interval
par(mfrow=2:1)
plot(x=newMTWRF$interval, y=newMTWRF$steps, 
     type="l", main="Avg Steps Per 5-Sec Interval",
     sub="Weekdays", xlab="Interval", ylab="Steps")

plot(x=newSS$interval, y=newSS$steps, type="l", 
     sub="Weekend", xlab="Interval", ylab="Steps")
```

This individual seems to be more active, in general terms, on the weekend than on weekdays.  They maintain a higher average number of steps on the weekend compared to weekdays for almost every 5 second interval. 
